### 2024-09-20 other/datasheet/datasheet_template.qmd
Add files via upload

---
title: "STA304 Paper 1: relationship between crime and mental health"
author: "Max LI"
date: "2024/9/20"
format: pdf
---

# Motivation

## Purpose
This dataset is created to explore the relationship between major crime indicators (MCI) and mental health apprehensions (MHA) across various districts in Toronto. The goal is to use crime data and MHA reports to predict crime trends using a linear regression model and identify any potential correlations between these factors.

## Context
The dataset combines two public datasets from Toronto's Open Data portal:
1. [Major Crime Indicators](https://open.toronto.ca/dataset/major-crime-indicators/)
2. [Mental Health Apprehensions](https://open.toronto.ca/dataset/mental-health-apprehensions/)

The analysis intends to assess whether MHA reports can explain variations in crime numbers and to identify residual patterns that may suggest areas for further investigation.

# Composition

## Variables
The dataset contains the following variables:\par
- **DIVISION**: The police division (or district) in Toronto, represented by codes such as D11, D12, etc.\par
- **num_crimes**: The total number of crimes recorded in a given division.\par
- **num_mha_reports**: The total number of MHA reports in the corresponding division.\par
- **predicted_crimes**: The predicted number of crimes in the division, generated from the linear model using MHA reports as a predictor.\par
- **residuals**: The difference between actual crime numbers and predicted values (actual - predicted), indicating how well the model performed for each division.\par

## Observations
The dataset includes 18 observations (one for each police division in Toronto).

## File Sources
- `raw_data_mha.csv`: Contains raw MHA reports by division.
- `raw_data_mci.csv`: Contains raw major crime reports by division.
- `clean_data.csv`: The combined and cleaned dataset with crime and MHA information.
- `clean_data_with_predictions.csv`: The final dataset that includes crime data, MHA reports, predictions, and residuals from the model.

# Data Sources

- **MHA Data**: [Mental Health Apprehensions](https://open.toronto.ca/dataset/mental-health-apprehensions/), provided by Toronto Police.
- **Crime Data**: [Major Crime Indicators](https://open.toronto.ca/dataset/major-crime-indicators/), provided by Toronto Police.
- **Model Outputs**: The final analysis dataset includes predictions from a linear regression model built in R.

# Collection Methods

1. **Download (01-download_data.R)**:
   The data was retrieved from Toronto’s open data portal and saved locally.

2. **Cleaning (02-clean_data.R)**:
   The MHA and MCI datasets were cleaned and merged by division. Each division's crime and MHA counts were aggregated to provide an overall view for the analysis.

3. **Testing (03-test_data.R)**:
   Initial checks and summaries were conducted to ensure that the cleaned data had valid and interpretable values.

4. **Modeling (04-model.R)**:
   A linear regression model was fit to predict `num_crimes` using `num_mha_reports` as the independent variable. The model output included predictions and residuals for each division.

5. **Replications (05-replications.R)**:
   The predictions from the model were compared to the actual crime data, and residuals were calculated. These were saved in a final CSV file.

# Limitations

- **Sample Size**: The dataset is limited to 18 divisions, which may not provide enough variability for a robust model.
- **Linear Model**: The assumption of a linear relationship between crime and MHA reports may not fully capture the complexities of these variables.
- **Residuals**: The residuals suggest that other variables (not captured in this dataset) may explain the difference between predicted and actual crime numbers.

# References

- Toronto Open Data Portal: Major Crime Indicators ([Link](https://open.toronto.ca/dataset/major-crime-indicators/))
- Toronto Open Data Portal: Mental Health Apprehensions ([Link](https://open.toronto.ca/dataset/mental-health-apprehensions/))

### 2024-09-20 other/literature/linear_model.qmd
Add files via upload

---
title: "Linear model"
author: "Max LI"
date: "2024/9/20"
format: pdf
---

## Model Summary

### Formula

The model predicts the number of crimes (`num_crimes`) based on the number of mental health apprehensions (`num_mha_reports`) using a linear relationship.

### Equation

$$
C = 6748.244 + 2.451 \times MHA
$$

### Coefficients and Interpretation

-   **C**\
    C stands for the predicted number of crimes, which is also the dependent variable in this model.

-   **Intercept:** 6748.244\
    This suggests that, even when there are no MHA reports, the baseline number of crimes in a division is approximately 6748.

-   **MHA:** 2.451\
    For each additional MHA report, the model predicts an increase of approximately 2.45 crimes in the corresponding division.

### Interpretation

The model suggests a positive linear relationship between MHA reports and crimes, meaning that as MHA reports increase, the number of crimes tends to increase as well.

### 2024-09-22 Paper/Paper.qmd
Add files via upload

---
title: "Mental health issue linked to major crime*"
author: Max LI
thanks: "Code and data are available at: https://github.com/ShaotongLi-Max/STA304-Paper-1-MaxLI."
date: 2024/9/21
date-format: long
abstract: "This Paper aims at discussing the relationship between Mental Health issue and committing crime. Using data from opendatatoronto, we can see Mental Health Act Apprehensions and major crime indicators of several police divisions in toronto. There appears to be positive linear relationship between the number of people having MHA apprehensions and number of crimes."
format: 
  pdf:
    toc: true           
    toc-title: "Table of Contents" 
    toc-depth: 2        
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(dplyr)
library(knitr)
library(gt)
library(broom)
```

# 1.Introduction

Mental Health Act is the law that describes what should happen when someone, in mental illness, should be treated to protect themselves and others around. The new Mental Health Act issued in 2014 set new regulations towards teenagers: "MHA Apprehensions of individuals aged 17 and under have been omitted to protect youth identity. From 2014 to 2020, these individuals comprised 6.5% (4,724 of 71,717) of all MHA Apprehensions, with individuals under 12 comprising 0.4% (320 of 71,717), and 12-17 comprising 6.1% (4,404 of 71,717) respectively. There are instances where an individual’s age group is classified as “Not Recorded”; these account for 1.3% (915 of 71,717) of all MHA Apprehensions." (*Mental Health Apprehensions*, 2023)[@mha_toronto] Now all data of MHA Apprehensions is stored in the data set provided by Toronto police department.

According to the research provided by Public Safety Canada, of the 5 million people who came into contact with police, approximately one in five (18.8%) met the criteria for a mental or substance use disorder. One in three Canadians with a mental health disorder reported having contact with police for at least one reason studied (34.4%).(*Annual Report on the State of Homelessness in Canada*,2017)[@PublicSafetyCanada] This is suggesting that there could be a positive relationship between Mental illness and crime rate.

This paper confirm that there is a positive linear relationship between mental health issue and major crime.First, it explores the data of MHA Apprehensions of all ages in several Toronto police divisions. Next, it explores the data of major crime numbers relating to each district. Finally, it develops a linear model of two data and plot the relationship between mental illness and major crime numbers.

# 2.Data

The data used in this paper came from the Open Data Toronto portal through the library opendatatoronto (Gelfand 2022)[@OpenDataToronto]. Data were cleaned and analyzed using the open source statistical programming language R (R Core Team 2023)[@R]. Libraries tidyverse (Wickham et al. 2019)[@tidyverse], dplyr (Wickham et al. 2023)[@dplyr], and lubridate (Grolemund and Wickham 2011)[@lubridate] were used for simulating, cleaning and testing. Graphics were made using ggplot2 (Wickham 2016)[@ggplot2], gt(Graham 2023)[@gt] and broom(Robinson and Wickham 2023)[@broom].

\newpage

# 3.1 Mental Health Act Apprehensions Data

The Mental Health Act Apprehensions Data are provided by Toronto Police Services. It includes all Mental Health Act (MHA) Apprehensions pursuant to the Mental Health Act, assigning different event numbers as distinct MHA Apprehensions. This data set is queried based on reported date and the last refreshed date is July 23,2024.

This data set specifically list the identity of the person who has MHA Apprehensions as well as the corresponding MHA type, which are as follows:

-   Section 15 (Form 1 - Physician - Application for Psychiatric Assessment);
-   Section 16 (Form 2 – Justice of the Peace – Order for Examination);
-   Section 17 (Police Officer’s Power of Apprehension);
-   Section 28 (1) (Form 9 - Elopee - Order for Return); and,
-   Section 33.4 (Form 47 - Community Treatment Order for Examination).

However, there are limitations to this data: A unique event number should not be interpreted as an apprehension of a distinct individual. An individual may have been apprehended multiple times under the Mental Health Act, with each apprehension recorded using a unique event number.(*Mental Health Apprehensions*,2023)[@mha_toronto]

Therefore, the columns chosen for this analysis were simply: "DIVISION" representing the Police Division where Offence Occurred. In cleaning data process, a new column "num_mha_reports" was made in the analysis data, representing the number of MHA reports in each police division.

# 3.2 Results of MHA Apprehensions Data

```{r,fig.pos='H'}
#| label: MHA Apprehensions Data
#| fig-cap: Number of MHA reports in each Police division (2014-2024)
#| echo: false
#| warning: false
#| message: false

data_file_path <- "/home/rstudio/STA304_Paper_1_MaxLI/data/analysis_data/clean_data.csv"
clean_data <- read.csv(data_file_path)

clean_data <- clean_data %>%
  arrange(DIVISION)

clean_data$DIVISION <- factor(clean_data$DIVISION, levels = clean_data$DIVISION)

ggplot(clean_data, aes(y = DIVISION, x = num_mha_reports)) +
  geom_bar(stat = "identity", fill = "#4E79A7", color = "black", width = 0.7) + # Custom color and bar width
  labs(title = "Number of MHA Reports by Division",
       y = "Police Division",
       x = "MHA Reports") +
  theme_minimal(base_size = 14) + # Improved theme with larger base font size
  theme(
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5), # Centered and bold title
    axis.text.y = element_text(size = 12), # Resized y-axis labels
    axis.title.y = element_text(face = "bold", size = 14),
    axis.title.x = element_text(face = "bold", size = 14)
  ) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0, max(clean_data$num_mha_reports) * 1.1)) + # Dynamic x-axis with padding
  geom_text(aes(label = num_mha_reports), hjust = -0.2, size = 4) # Display the values to the right of the bars

```

\newpage

# 4.1 Major crime indicator

The Major crime indicator data are provided by Toronto Police Services.It includes some major kinds of crime like Assault, Break and Enter, Auto Theft, Robbery and Theft Over that happened in every police division in Toronto. This data set is queried based on reported date and the last refreshed date is July 23,2024.

This data includes all MCI occurrences reported to the Toronto Police Service, including those where the location has not been able to be verified.However, this data does not include occurrences that have been deemed unfounded.(*Major Crime Indicators*,2023)[@mci_toronto]

The columns chosen for this analysis were simply: "DIVISION" representing the Police Division where major crime Occurred. In cleaning data process, a new column "num_crimes" was made in the analysis data, representing the number of major crimes took place in each police division.

# 4.2 Results of Major Crime Indicator Data

```{r,fig.pos='H'}
#| label: Major Crime Indicator Data
#| fig-cap: Number of Major Crime Indicator in each Police division (2014-2024)
#| echo: false
#| warning: false
#| message: false

abbreviate_number <- function(x) {
  ifelse(x >= 1000, paste0(round(x / 1000, 1), "k"), x)
}

ggplot(clean_data, aes(y = DIVISION, x = num_crimes)) +
  geom_bar(stat = "identity", fill = "#E15759", color = "black", width = 0.7) + 
  labs(title = "Number of Crimes by Division",
       y = "Division",
       x = "Crime Incidents") +
  theme_minimal(base_size = 14) + 
  theme(
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5), 
    axis.text.y = element_text(size = 12), 
    axis.title.y = element_text(face = "bold", size = 14),
    axis.title.x = element_text(face = "bold", size = 14)
  ) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0, max(clean_data$num_crimes) * 1.1), 
                     labels = abbreviate_number) + 
  geom_text(aes(label = abbreviate_number(num_crimes)), 
            hjust = -0.2, size = 4)

```

\newpage

# 5.MHA reports relationship to Major Crime Indicator

There appears to be a positive relationship between the number of MHA reports and the number of Major Crime Indicator. The higher the number of MHA reports, the more Major Crime Indicator in the same police division.

```{r,fig.pos='H'}
#| label: Linear model
#| fig-cap: Linear Model
#| echo: false
#| warning: false
#| message: false

clean_data <- read.csv("/home/rstudio/STA304_Paper_1_MaxLI/data/analysis_data/clean_data.csv")
model <- lm(num_crimes ~ num_mha_reports, data = clean_data)
par(mfrow = c(2, 2))
clean_data$predicted_crimes <- predict(model, newdata = clean_data)
ggplot(clean_data, aes(x = num_mha_reports, y = num_crimes)) +
  geom_point(color = "blue") +
  geom_line(aes(y = predicted_crimes), color = "red") +
  labs(title = "Number of Crimes vs. Number of MHA Reports",
       x = "Number of MHA Reports",
       y = "Number of Crimes") +
  theme_minimal()
```

Based on the positive relationship between crimes and MHA reports, a linear model is made to fit in this situation:

$$
C = 6748.244 + 2.451 \times MHA
$$ in which:

-   **C**\
    C stands for the predicted number of crimes, which is also the dependent variable in this model.

-   **Intercept:** 6748.244\
    This suggests that, even when there are no MHA reports, the baseline number of crimes in a division is approximately 6748.

-   **MHA:** 2.451\
    For each additional MHA report, the model predicts an increase of approximately 2.45 crimes in the corresponding division.

```{r,fig.pos='H'}
#| label: Linear model summary
#| fig-cap: The summary table for linear model
#| echo: false
#| warning: false
#| message: false

model_summary <- tidy(model)
model_stats <- glance(model)

coefficients_df <- model_summary %>%
  select(term, estimate, p.value)

model_stats_df <- data.frame(term = c("R-squared", "Adjusted R-squared"),
                              estimate = c(model_stats$r.squared, model_stats$adj.r.squared),
                              p.value = c(NA, NA))

final_summary_df <- coefficients_df  # Only keep coefficients_df

summary_table <- gt(data = final_summary_df) %>%
  tab_header(title = "Linear Model Summary") %>%
  text_transform(
    locations = cells_body(columns = vars(term)),
    fn = function(x) ifelse(x == "(Intercept)", "Intercept", x)
  ) %>%
  fmt_number(
    columns = c(estimate, p.value),  # Only format the estimate and p.value
    decimals = 3
  ) %>%
  cols_label(
    estimate = "Estimate",
    p.value = "P-value"
  ) %>%
  cols_align(align = "center", columns = vars(estimate, p.value)) %>%
  tab_footnote(
    footnote = paste("Adjusted R-squared:", round(model_stats$adj.r.squared, 3)),
    locations = cells_title(groups = "title")
  ) %>%
  tab_footnote(
    footnote = paste("R-squared:", round(model_stats$r.squared, 3)),
    locations = cells_title(groups = "title")
  )

summary_table

```

\newpage

# 6.Discussion

This paper investigated the relationship between the number of Mental Health Act Apprehensions and the major crime indicator, according to Toronto Police Services data retrieved from OpenData Toronto. Results shows that as the number of MHA reports rises, the major crime indicator also increases, based on data from 2014 to 2024. Furthermore, there appears to be a positive linear relationship between the number of MHA reports and the number of major crime indicator, which can be expressed as a linear model: C = 6748.244 + 2.451 \* MHA.

The result of this paper fits the result of the research by Public Safety Canada: "In the international research for both youth and adults analyzed for this report, findings suggest that certain mental health disorders, such as substance use disorder, conduct disorder, and antisocial personality disorder were shown to have a significant correlation to offending and re-offending." The p-value of our model is 4.19\*10\^-5, which is far blow the significance level of 0.05, indicating strong correlation between the number of Mental Health Act Apprehensions and the major crime indicator.(*Annual Report on the State of Homelessness in Canada*,2017)[@PublicSafetyCanada]

However, other variables like population could affect the number of both MHA reports and major crime indicator. Population is not included in Major crime indicator data and MHA Apprehensions Data. Adding population as a independent variable to the model could be a way of studying the relationship between MHA reports and major crime indicator.

The finding in this paper help to let people understand the link between mental health issue and crime. As further study or research confirm to the result of this paper, it could help government put more enphasis on enhancing health care for people with mental disorder and saving countless lives.

# 7.Limitation

The linear model indeed shows significant linear relationship between the number of MHA reports and major crime indicator. However, the adjusted R-squared in this model is 0.6388, indicating that there are 36.12% of the variance left unexplained by the model. This could imply that other factors, aside from MHA reports, might be influencing crime rates.

Also, the residual in this model is quite large. This could be due to the fact that there are only 20 police divisions in toronto, which is considered a small sample. From figure 4, the residual line is not a horizontal line, indicating the same problem.

```{r,fig.pos='H'}
#| label: Residual Plot
#| fig-cap: The Residual Vs Fitted value plot
#| echo: false
#| warning: false
#| message: false

plot(model, which = 1)
```

\newpage

# 8.Reference

R Core Team. 2023. R: A Language and Environment for Statistical Computing. Vienna, Austria: R Foundation for Statistical Computing. https://www.R-project.org/.

Gelfand, Jack. 2022. opendatatoronto: Accesses the City of Toronto Open Data Portal. R package version 0.1.0. https://CRAN.R-project.org/package=opendatatoronto.

Wickham, Hadley, Bryan, Jennifer, et al. 2019. Welcome to the Tidyverse. Journal of Open Source Software, 4(43), 1686. https://doi.org/10.21105/joss.01686.

Wickham, Hadley, François, Romain, Henry, Lionel, Müller, Kirill, and Vaughan, Davis. 2023. dplyr: A Grammar of Data Manipulation. R package version 1.1.2. https://CRAN.R-project.org/package=dplyr.

Grolemund, Garrett, and Wickham, Hadley. 2011. Dates and Times Made Easy with {lubridate}. Journal of Statistical Software, 40(3), 1-25. https://www.jstatsoft.org/article/view/v040i03.

Wickham, Hadley. 2016. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York. https://ggplot2.tidyverse.org.

Graham, J. 2023. gt: A Grammar of Tables. https://cran.r-project.org/web/packages/gt/index.html.

Robinson, D., and J. T. H. E. 2023. broom: Convert Statistical Analysis Objects into Tidy Data Frames. https://cran.r-project.org/web/packages/broom/index.html.

City of Toronto. 2022. Open Data Toronto Portal. https://opendata.toronto.ca/.

Public Safety Canada. 2017. 2017 Annual Report on the State of Homelessness in Canada. Government of Canada. https://www.publicsafety.gc.ca/cnt/rsrcs/pblctns/2017-h01-cp/index-en.aspx.

City of Toronto. 2023. Major Crime Indicators - Open Toronto. Toronto, Ontario: City of Toronto. https://open.toronto.ca/dataset/major-crime-indicators/.

City of Toronto. 2023. Mental Health Apprehensions - Open Toronto. Toronto, Ontario: City of Toronto. https://open.toronto.ca/dataset/mental-health-apprehensions/.

### 2024-09-22 Paper/Paper.qmd
Add files via upload

@@
 ---
-title: "Mental health issue linked to major crime*"
+title: "Mental health issue linked to major crime"
 author: Max LI
@@
 ggplot(clean_data, aes(y = DIVISION, x = num_mha_reports)) +
-  geom_bar(stat = "identity", fill = "#4E79A7", color = "black", width = 0.7) + # Custom color and bar width
+  geom_bar(stat = "identity", fill = "#4E79A7", color = "black", width = 0.7, position = position_dodge(width = 0.9)) +  
   labs(title = "Number of MHA Reports by Division",
@@
        x = "MHA Reports") +
-  theme_minimal(base_size = 14) + # Improved theme with larger base font size
+  theme_minimal(base_size = 14) + 
   theme(
-    plot.title = element_text(face = "bold", size = 16, hjust = 0.5), # Centered and bold title
-    axis.text.y = element_text(size = 12), # Resized y-axis labels
+    plot.title = element_text(face = "bold", size = 16, hjust = 0.5), 
+    axis.text.y = element_text(size = 12),  
     axis.title.y = element_text(face = "bold", size = 14),
@@
   ) +
-  scale_x_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0, max(clean_data$num_mha_reports) * 1.1)) + # Dynamic x-axis with padding
-  geom_text(aes(label = num_mha_reports), hjust = -0.2, size = 4) # Display the values to the right of the bars
+  scale_x_continuous(expand = expansion(mult = c(0, 0.1)), limits = c(0, max(clean_data$num_mha_reports) * 1.05)) +  # Adjust the padding on the x-axis
+  geom_text(aes(label = num_mha_reports), hjust = -0.3, size = 3.5)
 
@@
   ) +
-  scale_x_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0, max(clean_data$num_crimes) * 1.1), 
+  scale_x_continuous(expand = expansion(mult = c(0, 0.1)), limits = c(0, max(clean_data$num_crimes) * 1.05), 
                      labels = abbreviate_number) +  
   geom_text(aes(label = abbreviate_number(num_crimes)), 
-            hjust = -0.2, size = 4)
+            hjust = -0.3, size = 3.5)
 
@@
 #| label: Linear model summary
-#| fig-cap: The summary table for linear model
 #| echo: false
@@
 
-final_summary_df <- coefficients_df  # Only keep coefficients_df
+final_summary_df <- coefficients_df
 
@@
   fmt_number(
-    columns = c(estimate, p.value),  # Only format the estimate and p.value
+    columns = c(estimate, p.value),  
     decimals = 3
@@
   tab_footnote(
+    footnote = paste("R-squared:", round(model_stats$r.squared, 3)),
+    locations = cells_title(groups = "title")
+  ) %>%
+  tab_footnote(
     footnote = paste("Adjusted R-squared:", round(model_stats$adj.r.squared, 3)),
@@
   ) %>%
-  tab_footnote(
-    footnote = paste("R-squared:", round(model_stats$r.squared, 3)),
-    locations = cells_title(groups = "title")
-  )
+  opt_footnote_marks(marks = "letters")
 

### 2024-09-25 Paper/Paper.qmd
Add files via upload

@@
 date-format: long
-abstract: "This Paper aims at discussing the relationship between Mental Health issue and committing crime.
-Using data from opendatatoronto, we can see Mental Health Act Apprehensions and major crime indicators of several police divisions in toronto.
-There appears to be positive linear relationship between the number of people having MHA apprehensions and number of crimes."
+abstract: "This paper discusses the relationship between mental health issues and crime, focusing on Mental Health Act Apprehensions and major crime indicators in several Toronto police divisions.
+A positive linear relationship is found between the number of MHA apprehensions and reported crimes, suggesting that mental health-related interventions could help reduce crime rates while supporting affected individuals."
 format: 
@@
     toc-depth: 2    
+    number_sections: true    
 bibliography: references.bib
@@
 From 2014 to 2020, these individuals comprised 6.5% (4,724 of 71,717) of all MHA Apprehensions, with individuals under 12 comprising 0.4% (320 of 71,717), and 12-17 comprising 6.1% (4,404 of 71,717) respectively.
-There are instances where an individual’s age group is classified as “Not Recorded”; these account for 1.3% (915 of 71,717) of all MHA Apprehensions." (*Mental Health Apprehensions*, 2023)[@mha_toronto] Now all data of MHA Apprehensions is stored in the data set provided by Toronto police department.
+There are instances where an individual’s age group is classified as “Not Recorded”; these account for 1.3% (915 of 71,717) of all MHA Apprehensions."[@mha_toronto] Now all data of MHA Apprehensions is stored in the data set provided by Toronto police department.
 
 According to the research provided by Public Safety Canada, of the 5 million people who came into contact with police, approximately one in five (18.8%) met the criteria for a mental or substance use disorder.
-One in three Canadians with a mental health disorder reported having contact with police for at least one reason studied (34.4%).(*Annual Report on the State of Homelessness in Canada*,2017)[@PublicSafetyCanada] This is suggesting that there could be a positive relationship between Mental illness and crime rate.
+One in three Canadians with a mental health disorder reported having contact with police for at least one reason studied (34.4%).[@PublicSafetyCanada] This is suggesting that there could be a positive relationship between Mental illness and crime rate.
 
-This paper confirm that there is a positive linear relationship between mental health issue and major crime.First, it explores the data of MHA Apprehensions of all ages in several Toronto police divisions.
-Next, it explores the data of major crime numbers relating to each district.
-Finally, it develops a linear model of two data and plot the relationship between mental illness and major crime numbers.
+This paper investigates the relationship between mental health apprehensions under the Mental Health Act (MHA) and major crime incidents across various police divisions in Toronto.
+Using two datasets, this paper analyzes the distribution of both mental health-related incidents and crimes across divisions through visualizations and statistical methods.It further develops a linear regression model to quantify and explore the positive correlation between MHA reports and crime rates, suggesting that a better understanding of this relationship could aid in crime control and mental health care strategies.
+
+In section 3 and 4, this paper shows descriptions and visualizations of MHA apprehensions data as well as major crimes across different police divisions, with plots depict how the number of two factors varies across divisions, providing insights into the distribution of mental health-related events.
+
+In section 5, this paper compares MHA apprehensions with crime data and develops the positive linear regression model between the two factors.
+In section 6, the paper further discuss the model and how it can be used in reality.
+
+In section 7, this paper explains the limitation of the original data and the model.
+
+\newpage
 
@@
 
-The data used in this paper came from the Open Data Toronto portal through the library opendatatoronto (Gelfand 2022)[@OpenDataToronto].
-Data were cleaned and analyzed using the open source statistical programming language R (R Core Team 2023)[@R].
-Libraries tidyverse (Wickham et al.
-2019)[@tidyverse], dplyr (Wickham et al.
-2023)[@dplyr], and lubridate (Grolemund and Wickham 2011)[@lubridate] were used for simulating, cleaning and testing.
-Graphics were made using ggplot2 (Wickham 2016)[@ggplot2], gt(Graham 2023)[@gt] and broom(Robinson and Wickham 2023)[@broom].
+The data used in this paper came from the Open Data Toronto portal through the library opendatatoronto [@OpenDataToronto].
+Data were cleaned and analyzed using the open source statistical programming language R [@R].
+Libraries tidyverse [@tidyverse], dplyr [@dplyr], and lubridate [@lubridate] were used for simulating, cleaning and testing.
+Graphics were made using ggplot2 [@ggplot2], gt[@gt] and broom[@broom].
+
+The cleaned data comes from two raw data sets: Mental Health Act Apprehensions Dataset, which contains records of Mental Health Act (MHA) reports grouped by police division, and Major Crime Indicators Dataset, which includes records of crimes similarly categorized by police division.
+The cleaned data set aggregates the number of MHA reports and crimes for each district, providing a comprehensive overview of the relationship between mental health reports and crime incidents across the specified districts.
 
@@
 
-# 3.1 Mental Health Act Apprehensions Data
+# 3.Mental Health Act Apprehensions Data
 
@@
 However, there are limitations to this data: A unique event number should not be interpreted as an apprehension of a distinct individual.
-An individual may have been apprehended multiple times under the Mental Health Act, with each apprehension recorded using a unique event number.(*Mental Health Apprehensions*,2023)[@mha_toronto]
+An individual may have been apprehended multiple times under the Mental Health Act, with each apprehension recorded using a unique event number.[@mha_toronto]
 
@@
 
-# 3.2 Results of MHA Apprehensions Data
+\newpage
+
+## 3.1 Results of MHA Apprehensions Data
 
@@
   theme(
-    plot.title = element_text(face = "bold", size = 16, hjust = 0.5), 
-    axis.text.y = element_text(size = 12),  
-    axis.title.y = element_text(face = "bold", size = 14),
-    axis.title.x = element_text(face = "bold", size = 14)
+    plot.title = element_text(face = "bold", size = 14, hjust = 0.5), 
+    axis.text.y = element_text(size = 10),  
+    axis.title.y = element_text(face = "bold", size = 12),
+    axis.title.x = element_text(face = "bold", size = 12)
   ) +
-  scale_x_continuous(expand = expansion(mult = c(0, 0.1)), limits = c(0, max(clean_data$num_mha_reports) * 1.05)) +  # Adjust the padding on the x-axis
+  scale_x_continuous(expand = expansion(mult = c(0, 0.1)), limits = c(0, max(clean_data$num_mha_reports) * 1.05)) +  
   geom_text(aes(label = num_mha_reports), hjust = -0.3, size = 3.5)
+
+```
+
+```{r,fig.pos='H'}
+#| label: MHA Apprehensions Data 2
+#| fig-cap: Number of MHA reports in each Police division (2014-2024)
+#| echo: false
+#| warning: false
+#| message: false
+
+ggplot(clean_data, aes(x = DIVISION, y = num_mha_reports, group = 1)) +
+  geom_line(color = "#4E79A7", size = 1) +
+  geom_point(size = 3, color = "#4E79A7") + 
+  labs(title = "Number of MHA Reports by Division",
+       x = "Police Division", 
+       y = "MHA Reports") +    
+  theme_minimal(base_size = 14) + 
+  theme(
+    plot.title = element_text(face = "bold", size = 14, hjust = 0.5), 
+    axis.text.x = element_text(size = 10, angle = 45, hjust = 1),  
+    axis.title.y = element_text(face = "bold", size = 12),
+    axis.title.x = element_text(face = "bold", size = 12)
+  ) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.1)), limits = c(0, max(clean_data$num_mha_reports) * 1.05)) + 
+  geom_text(aes(label = num_mha_reports), vjust = -0.5, size = 3.5)
 
@@
 
-# 4.1 Major crime indicator
+# 4.Major crime indicator
 
-The Major crime indicator data are provided by Toronto Police Services.It includes some major kinds of crime like Assault, Break and Enter, Auto Theft, Robbery and Theft Over that happened in every police division in Toronto.
-This data set is queried based on reported date and the last refreshed date is July 23,2024.
+This data set includes all Major Crime Indicators (MCI) occurrences by reported date.
+The Major Crime Indicators categories are Assault, Break and Enter, Auto Theft, Robbery and Theft Over (Excludes Sexual Assaults).[@mci_toronto]This data set is queried based on reported date and the last refreshed date is July 23,2024.
 
-This data includes all MCI occurrences reported to the Toronto Police Service, including those where the location has not been able to be verified.However, this data does not include occurrences that have been deemed unfounded.(*Major Crime Indicators*,2023)[@mci_toronto]
+This data includes all MCI occurrences reported to the Toronto Police Service, including those where the location has not been able to be verified.This is not an issue since this paper only takes into account the number of crimes in different division.This data does not include occurrences that have been deemed unfounded, or "has been determined through police investigation that the offence reported did not occur, nor was it attempted".Therefore, every crime in the dataset is a crime that has been confirmed by the police department.[@mci_toronto]
 
@@
 
-# 4.2 Results of Major Crime Indicator Data
+\newpage
+
+## 4.1 Results of Major Crime Indicator Data
 
@@
   theme(
-    plot.title = element_text(face = "bold", size = 16, hjust = 0.5),  
-    axis.text.y = element_text(size = 12), 
-    axis.title.y = element_text(face = "bold", size = 14),
-    axis.title.x = element_text(face = "bold", size = 14)
+    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),  
+    axis.text.y = element_text(size = 10), 
+    axis.title.y = element_text(face = "bold", size = 12),
+    axis.title.x = element_text(face = "bold", size = 12)
   ) +
@@
 
+```{r,fig.pos='H'}
+#| label: Major Crime Indicator Data 2
+#| fig-cap: Number of Major Crime Indicator in each Police division (2014-2024)
+#| echo: false
+#| warning: false
+#| message: false
+
+ggplot(clean_data, aes(x = DIVISION, y = num_crimes, group = 1)) +
+  geom_line(color = "#E15759", size = 1) +
+  geom_point(size = 3, color = "#E15759") + 
+  labs(title = "Number of Crimes by Division",
+       x = "Division",  
+       y = "Crime Incidents") +  
+  theme_minimal(base_size = 14) + 
+  theme(
+    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),  
+    axis.text.x = element_text(size = 10, angle = 45, hjust = 1), 
+    axis.title.y = element_text(face = "bold", size = 12),
+    axis.title.x = element_text(face = "bold", size = 12)
+  ) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.1)), 
+                     limits = c(0, max(clean_data$num_crimes) * 1.05), 
+                     labels = abbreviate_number) +  
+  geom_text(aes(label = abbreviate_number(num_crimes)), 
+            vjust = -0.5, size = 3.5)
+
+```
+
 \newpage
@@
 
-There appears to be a positive relationship between the number of MHA reports and the number of Major Crime Indicator.
-The higher the number of MHA reports, the more Major Crime Indicator in the same police division.
-
 ```{r,fig.pos='H'}
-#| label: Linear model
-#| fig-cap: Linear Model
+#| label: comparison 
+#| fig-cap: comparison between two data
 #| echo: false
@@
 
-clean_data <- read.csv("/home/rstudio/STA304_Paper_1_MaxLI/data/analysis_data/clean_data.csv")
-model <- lm(num_crimes ~ num_mha_reports, data = clean_data)
-par(mfrow = c(2, 2))
-clean_data$predicted_crimes <- predict(model, newdata = clean_data)
-ggplot(clean_data, aes(x = num_mha_reports, y = num_crimes)) +
-  geom_point(color = "blue") +
-  geom_line(aes(y = predicted_crimes), color = "red") +
-  labs(title = "Number of Crimes vs. Number of MHA Reports",
-       x = "Number of MHA Reports",
-       y = "Number of Crimes") +
-  theme_minimal()
+scale_factor <- max(clean_data$num_mha_reports) / max(clean_data$num_crimes)
+ggplot(clean_data, aes(y = DIVISION)) +
+  geom_bar(aes(x = num_mha_reports), stat = "identity", fill = "#4E79A7", color = "black", width = 0.5) +
+  geom_bar(aes(x = num_crimes * scale_factor), stat = "identity", fill = "#E15759", color = "black", width = 0.5, position = position_nudge(y = -0.5)) +
+  labs(title = "Comparison of MHA Reports and Crimes by Division",
+       y = "Division",
+       x = "MHA Reports / Crime Incidents") +
+  theme_minimal(base_size = 16) + 
+  theme(
+    plot.title = element_text(face = "bold", size = 12, hjust = 0.5),  
+    axis.text.y = element_text(size = 10),  
+    axis.title.y = element_text(face = "bold", size = 12),
+    axis.title.x = element_text(face = "bold", size = 12)
+  ) +
+  scale_x_continuous(sec.axis = sec_axis(~./scale_factor, name = "Crime Incidents"))
+
 ```
 
-Based on the positive relationship between crimes and MHA reports, a linear model is made to fit in this situation:
+```{r,fig.pos='H'}
+#| label: comparison 2
+#| fig-cap: comparison between two data
+#| echo: false
+#| warning: false
+#| message: false
+
+scale_factor <- max(clean_data$num_mha_reports) / max(clean_data$num_crimes)
+
+ggplot(clean_data, aes(x = DIVISION)) +
+  geom_line(aes(y = num_mha_reports, group = 1), color = "#4E79A7", size = 1) +  
+  geom_point(aes(y = num_mha_reports), size = 3, color = "#4E79A7") +
+  geom_line(aes(y = num_crimes * scale_factor, group = 1), color = "#E15759", size = 1) +  
+  geom_point(aes(y = num_crimes * scale_factor), size = 3, color = "#E15759") +
+  labs(title = "Comparison of MHA Reports and Crimes by Division",
+       x = "Division",
+       y = "MHA Reports") +
+  theme_minimal(base_size = 14) + 
+  theme(
+    plot.title = element_text(face = "bold", size = 12, hjust = 0.5),  
+    axis.text.x = element_text(size = 10, angle = 45, hjust = 1), 
+    axis.title.y = element_text(face = "bold", size = 12),
+    axis.title.x = element_text(face = "bold", size = 12)
+  ) +
+  scale_y_continuous(sec.axis = sec_axis(~./scale_factor, name = "Crime Incidents", labels = abbreviate_number))
+
+```
+
+By comparing the two line plot,a positive relationship between the number of MHA reports and the number of Major Crime Indicator, the number of MHA reports seems to influence the number of crime incidents accross different police divisions.
+In order to make sure there are linear relationship between two factors, a linear model is made to fit in this situation based on two data:
 
@@
     For each additional MHA report, the model predicts an increase of approximately 2.45 crimes in the corresponding division.
+
+The corresponding graph of linear model is:
+
+```{r,fig.pos='H'}
+#| label: Linear model
+#| fig-cap: Linear Model
+#| echo: false
+#| warning: false
+#| message: false
+
+model <- lm(num_crimes ~ num_mha_reports, data = clean_data)
+clean_data$predicted_crimes <- predict(model, newdata = clean_data)
+ggplot(clean_data, aes(x = num_mha_reports, y = num_crimes)) +
+  geom_point(color = "blue") +  
+  geom_line(aes(y = predicted_crimes), color = "red") + 
+  labs(title = "Number of Crimes vs. Number of MHA Reports",
+       x = "Number of MHA Reports",
+       y = "Number of Crimes") +
+  theme_minimal()
+```
+
+The p-value for both intercept and coefficient (MHA Reports) are below the significant level of 0.05, indicating a strong evidence that the relationship between MHA Reports and major crime indicator follows this linear model.
 
@@
   ) %>%
-  cols_align(align = "center", columns = vars(estimate, p.value)) %>%
-  tab_footnote(
-    footnote = paste("R-squared:", round(model_stats$r.squared, 3)),
-    locations = cells_title(groups = "title")
-  ) %>%
-  tab_footnote(
-    footnote = paste("Adjusted R-squared:", round(model_stats$adj.r.squared, 3)),
-    locations = cells_title(groups = "title")
-  ) %>%
-  opt_footnote_marks(marks = "letters")
+  cols_align(align = "center", columns = vars(estimate, p.value))
 
@@
 
-The result of this paper fits the result of the research by Public Safety Canada: "In the international research for both youth and adults analyzed for this report, findings suggest that certain mental health disorders, such as substance use disorder, conduct disorder, and antisocial personality disorder were shown to have a significant correlation to offending and re-offending." The p-value of our model is 4.19\*10\^-5, which is far blow the significance level of 0.05, indicating strong correlation between the number of Mental Health Act Apprehensions and the major crime indicator.(*Annual Report on the State of Homelessness in Canada*,2017)[@PublicSafetyCanada]
+The result of this paper shares the insight of the research by Public Safety Canada: "In the international research for both youth and adults analyzed for this report, findings suggest that certain mental health disorders, such as substance use disorder, conduct disorder, and antisocial personality disorder were shown to have a significant correlation to offending and re-offending." The p-value of our model is 4.19\*10\^-5, which is far blow the significance level of 0.05, indicating strong correlation between the number of Mental Health Act Apprehensions and the major crime indicator.[@PublicSafetyCanada]
 
@@
 
-The finding in this paper help to let people understand the link between mental health issue and crime.
-As further study or research confirm to the result of this paper, it could help government put more enphasis on enhancing health care for people with mental disorder and saving countless lives.
+This paper and its model can help the government and society better allocate resources for both mental health services and law enforcement.
+By understanding the correlation between mental health apprehensions and crime rates, policymakers can implement targeted interventions to address underlying issues, such as providing mental health support in high-crime areas or improving police training on handling mental health incidents.
+Additionally, this model can assist in predicting areas where both mental health-related incidents and crime are likely to occur, helping authorities to proactively implement prevention strategies, potentially reducing crime and improving mental health outcomes.
+
+\newpage
 
 # 7.Limitation
+The limitation of this paper is split into two parts: the limitation of data and the limitation of linear model
 
+## 7.1 Data outlier
+Significant outliers presents in both MHA reports and Major Crime Indicator data, which is possibly caused by a relative small smaple population as there are only 20 police division in Toronto.
+
+```{r,fig.pos='H'}
+#| label: Box Plot
+#| fig-cap: The box plot of two factors
+#| echo: false
+#| warning: false
+#| message: false
+
+data_file_path <- "/home/rstudio/STA304_Paper_1_MaxLI/data/analysis_data/clean_data.csv"
+clean_data <- read.csv(data_file_path)
+
+long_data <- clean_data %>%
+  select(DIVISION, num_mha_reports, num_crimes) %>%
+  pivot_longer(cols = c(num_mha_reports, num_crimes),
+               names_to = "Indicator",
+               values_to = "Count")
+
+ggplot(long_data, aes(x = Indicator, y = Count)) +
+  geom_boxplot(outlier.colour = "black", outlier.shape = 16, outlier.size = 2, fill = NA, color = "black") +
+  geom_jitter(width = 0.2, alpha = 0.5) +  
+  labs(title = "Boxplot of MHA Reports and Crimes",
+       x = "Indicator",
+       y = "Count") +
+  theme_minimal(base_size = 15) +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 15),  
+    axis.title.x = element_text(hjust = 0.5, size = 15),
+    axis.title.y = element_text(hjust = 0.5, size = 15)  
+  )
+```
+
+\newpage
+
+## 7.2 Linear model
 The linear model indeed shows significant linear relationship between the number of MHA reports and major crime indicator.
@@
 
-Also, the residual in this model is quite large.
-This could be due to the fact that there are only 20 police divisions in toronto, which is considered a small sample.
-From figure 4, the residual line is not a horizontal line, indicating the same problem.
+```{r,fig.pos='H'}
+#| label: Summary table
+#| fig-cap: Summary table of linear model
+#| echo: false
+#| warning: false
+#| message: false
+
+model_summary <- tidy(model)
+model_stats <- glance(model)
+summary_stats <- tibble(
+  Statistic = c("Residual Standard Error", "Multiple R-squared", "Adjusted R-squared", "F-statistic", "p-value"),
+  Value = c(round(model_stats$sigma, 3), 
+            round(model_stats$r.squared, 3), 
+            round(model_stats$adj.r.squared, 3), 
+            round(model_stats$statistic, 2), 
+            formatC(model_stats$p.value, format = "e", digits = 2))
+)
+
+summary_table <- summary_stats %>%
+  gt() %>%
+  tab_header(title = "Linear Model Summary") %>%
+  fmt_number(
+    columns = vars(Value),
+    decimals = 3
+  ) %>%
+  cols_label(
+    Statistic = "Statistic",
+    Value = "Value"
+  ) %>%
+  cols_align(align = "center", columns = vars(Statistic, Value))
+summary_table
+```
+
+## 7.3 Model residual
+The residuals in this model are quite large, leading to a complex residual pattern, which may indicate potential issues with model fit or the presence of unaccounted variability in the data.
 
@@
 
-plot(model, which = 1)
+plot(model, which=1)
 ```
@@
 # 8.Reference
-
-R Core Team. 2023. R: A Language and Environment for Statistical Computing. Vienna, Austria: R Foundation for Statistical Computing. https://www.R-project.org/.
-
-Gelfand, Jack. 2022. opendatatoronto: Accesses the City of Toronto Open Data Portal. R package version 0.1.0. https://CRAN.R-project.org/package=opendatatoronto.
-
-Wickham, Hadley, Bryan, Jennifer, et al. 2019. Welcome to the Tidyverse. Journal of Open Source Software, 4(43), 1686. https://doi.org/10.21105/joss.01686.
-
-Wickham, Hadley, François, Romain, Henry, Lionel, Müller, Kirill, and Vaughan, Davis. 2023. dplyr: A Grammar of Data Manipulation. R package version 1.1.2. https://CRAN.R-project.org/package=dplyr.
-
-Grolemund, Garrett, and Wickham, Hadley. 2011. Dates and Times Made Easy with {lubridate}. Journal of Statistical Software, 40(3), 1-25. https://www.jstatsoft.org/article/view/v040i03.
-
-Wickham, Hadley. 2016. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York. https://ggplot2.tidyverse.org.
-
-Graham, J. 2023. gt: A Grammar of Tables. https://cran.r-project.org/web/packages/gt/index.html.
-
-Robinson, D., and J. T. H. E. 2023. broom: Convert Statistical Analysis Objects into Tidy Data Frames. https://cran.r-project.org/web/packages/broom/index.html.
-
-City of Toronto. 2022. Open Data Toronto Portal. https://opendata.toronto.ca/.
-
-Public Safety Canada. 2017. 2017 Annual Report on the State of Homelessness in Canada. Government of Canada. https://www.publicsafety.gc.ca/cnt/rsrcs/pblctns/2017-h01-cp/index-en.aspx.
-
-City of Toronto. 2023. Major Crime Indicators - Open Toronto. Toronto, Ontario: City of Toronto. https://open.toronto.ca/dataset/major-crime-indicators/.
-
-City of Toronto. 2023. Mental Health Apprehensions - Open Toronto. Toronto, Ontario: City of Toronto. https://open.toronto.ca/dataset/mental-health-apprehensions/.

### 2024-09-25 Paper/Paper.qmd
Add files via upload

@@
 ---
-title: "Mental health issue linked to major crime"
+title: "Mental health issue linked to major crimes"
 author: Max LI
-thanks: "Code and data are available at: https://github.com/ShaotongLi-Max/STA304-Paper-1-MaxLI."
+thanks: "Code and data are available at: https://github.com/ShaotongLi-Max/Mental_health_issue_related_to_major_crime."
 date: 2024/9/21
@@
 abstract: "This paper discusses the relationship between mental health issues and crime, focusing on Mental Health Act Apprehensions and major crime indicators in several Toronto police divisions.
-A positive linear relationship is found between the number of MHA apprehensions and reported crimes, suggesting that mental health-related interventions could help reduce crime rates while supporting affected individuals."
+A positive linear relationship is found between the number of MHA apprehensions and reported crimes, suggesting that mental health-related interventions could help reduce crime rates while also supporting affected individuals."
 format: 
@@
 
-Mental Health Act is the law that describes what should happen when someone, in mental illness, should be treated to protect themselves and others around.
+Mental Health Act is a law that describes what should happen when someone with a mental illness should be treated to protect themselves and others around.
 The new Mental Health Act issued in 2014 set new regulations towards teenagers: "MHA Apprehensions of individuals aged 17 and under have been omitted to protect youth identity.
@@
 According to the research provided by Public Safety Canada, of the 5 million people who came into contact with police, approximately one in five (18.8%) met the criteria for a mental or substance use disorder.
-One in three Canadians with a mental health disorder reported having contact with police for at least one reason studied (34.4%).[@PublicSafetyCanada] This is suggesting that there could be a positive relationship between Mental illness and crime rate.
+One in three Canadians with a mental health disorder reported having contact with police for at least one reason studied (34.4%).[@PublicSafetyCanada] This is suggesting that there could be a positive relationship between mental illness and crime rates.
 
@@
 
-In section 7, this paper explains the limitation of the original data and the model.
+In section 7, this paper explains the limitation of the original data and the model. In section 8, this paper lists the cleaned data as well as the full model summary table.
 
@@
 ggplot(clean_data, aes(y = DIVISION, x = num_mha_reports)) +
-  geom_bar(stat = "identity", fill = "#4E79A7", color = "black", width = 0.7, position = position_dodge(width = 0.9)) +  
+  geom_bar(stat = "identity", fill = "#A9CCE3", color = "#A9CCE3", width = 0.7, position = position_dodge(width = 0.9)) +  
   labs(title = "Number of MHA Reports by Division",
@@
 ggplot(clean_data, aes(x = DIVISION, y = num_mha_reports, group = 1)) +
-  geom_line(color = "#4E79A7", size = 1) +
-  geom_point(size = 3, color = "#4E79A7") + 
+  geom_line(color = "#A9CCE3", size = 1) +
+  geom_point(size = 3, color = "#A9CCE3") + 
   labs(title = "Number of MHA Reports by Division",
@@
 ggplot(clean_data, aes(y = DIVISION, x = num_crimes)) +
-  geom_bar(stat = "identity", fill = "#E15759", color = "black", width = 0.7) +  
+  geom_bar(stat = "identity", fill = "#F1948A", color = "#F1948A", width = 0.7) +  
   labs(title = "Number of Crimes by Division",
@@
 ggplot(clean_data, aes(x = DIVISION, y = num_crimes, group = 1)) +
-  geom_line(color = "#E15759", size = 1) +
-  geom_point(size = 3, color = "#E15759") + 
+  geom_line(color = "#F1948A", size = 1) +
+  geom_point(size = 3, color = "#F1948A") + 
   labs(title = "Number of Crimes by Division",
@@
 ggplot(clean_data, aes(y = DIVISION)) +
-  geom_bar(aes(x = num_mha_reports), stat = "identity", fill = "#4E79A7", color = "black", width = 0.5) +
-  geom_bar(aes(x = num_crimes * scale_factor), stat = "identity", fill = "#E15759", color = "black", width = 0.5, position = position_nudge(y = -0.5)) +
+  geom_bar(aes(x = num_mha_reports), stat = "identity", fill = "#A9CCE3", color = "#A9CCE3", width = 0.5) +
+  geom_bar(aes(x = num_crimes * scale_factor), stat = "identity", fill = "#F1948A", color = "#F1948A", width = 0.5, position = position_nudge(y = -0.5)) +
   labs(title = "Comparison of MHA Reports and Crimes by Division",
@@
 ggplot(clean_data, aes(x = DIVISION)) +
-  geom_line(aes(y = num_mha_reports, group = 1), color = "#4E79A7", size = 1) +  
-  geom_point(aes(y = num_mha_reports), size = 3, color = "#4E79A7") +
-  geom_line(aes(y = num_crimes * scale_factor, group = 1), color = "#E15759", size = 1) +  
-  geom_point(aes(y = num_crimes * scale_factor), size = 3, color = "#E15759") +
+  geom_line(aes(y = num_mha_reports, group = 1), color = "#A9CCE3", size = 1) +  
+  geom_point(aes(y = num_mha_reports), size = 3, color = "#A9CCE3") +
+  geom_line(aes(y = num_crimes * scale_factor, group = 1), color = "#F1948A", size = 1) +  
+  geom_point(aes(y = num_crimes * scale_factor), size = 3, color = "#F1948A") +
   labs(title = "Comparison of MHA Reports and Crimes by Division",
@@
 
-By comparing the two line plot,a positive relationship between the number of MHA reports and the number of Major Crime Indicator, the number of MHA reports seems to influence the number of crime incidents accross different police divisions.
-In order to make sure there are linear relationship between two factors, a linear model is made to fit in this situation based on two data:
+By comparing the two line plots,a positive relationship between the number of MHA reports and the number of Major Crime Indicator, the number of MHA reports seems to influence the number of crime incidents accross different police divisions.
+In order to confirm there are linear relationship between two factors, a linear model is made to fit in this situation based on two data:
 
@@
 ggplot(clean_data, aes(x = num_mha_reports, y = num_crimes)) +
-  geom_point(color = "blue") +  
-  geom_line(aes(y = predicted_crimes), color = "red") + 
+  geom_point(color = "#A9CCE3") +  
+  geom_line(aes(y = predicted_crimes), color = "#F1948A") + 
   labs(title = "Number of Crimes vs. Number of MHA Reports",
@@
 
-The p-value for both intercept and coefficient (MHA Reports) are below the significant level of 0.05, indicating a strong evidence that the relationship between MHA Reports and major crime indicator follows this linear model.
+The p-values for both intercept and coefficient (MHA Reports) are below the significant level of 0.05, indicating a strong evidence that the relationship between MHA Reports and major crime indicator follows this linear model.
 
@@
 summary_table <- gt(data = final_summary_df) %>%
-  tab_header(title = "Linear Model Summary") %>%
+  tab_header(title = "Linear Model Summary 1") %>%
   text_transform(
@@
 
-The result of this paper shares the insight of the research by Public Safety Canada: "In the international research for both youth and adults analyzed for this report, findings suggest that certain mental health disorders, such as substance use disorder, conduct disorder, and antisocial personality disorder were shown to have a significant correlation to offending and re-offending." The p-value of our model is 4.19\*10\^-5, which is far blow the significance level of 0.05, indicating strong correlation between the number of Mental Health Act Apprehensions and the major crime indicator.[@PublicSafetyCanada]
+The result of this paper shares the insight of the research by Public Safety Canada: "In the international research for both youth and adults analyzed for this report, findings suggest that certain mental health disorders, such as substance use disorder, conduct disorder, and antisocial personality disorder were shown to have a significant correlation to offending and re-offending." The p-value of our model is 4.19\*10\^-5, which is far below the significance level of 0.05, indicating a strong correlation between the number of Mental Health Act Apprehensions and the major crime indicator.[@PublicSafetyCanada]
 
@@
 Population is not included in Major crime indicator data and MHA Apprehensions Data.
-Adding population as a independent variable to the model could be a way of studying the relationship between MHA reports and major crime indicator.
+Adding population as an independent variable to the model could be a way of studying the relationship between MHA reports and major crime indicator.
 
-This paper and its model can help the government and society better allocate resources for both mental health services and law enforcement.
+This paper and its model can assist the government and society better allocate resources for both mental health services and law enforcement.
 By understanding the correlation between mental health apprehensions and crime rates, policymakers can implement targeted interventions to address underlying issues, such as providing mental health support in high-crime areas or improving police training on handling mental health incidents.
-Additionally, this model can assist in predicting areas where both mental health-related incidents and crime are likely to occur, helping authorities to proactively implement prevention strategies, potentially reducing crime and improving mental health outcomes.
+Additionally, this model can assist in predicting areas where both mental health-related incidents and crime are likely to occur, helping authorities proactively implement prevention strategies, which could potentially reducing crime and improving mental health outcomes.
 
@@
 # 7.Limitation
-The limitation of this paper is split into two parts: the limitation of data and the limitation of linear model
+
+The limitation of this paper is split into two parts: the limitation of data (7.1) and the limitation of linear model (7.2 & 7.3)
 
 ## 7.1 Data outlier
-Significant outliers presents in both MHA reports and Major Crime Indicator data, which is possibly caused by a relative small smaple population as there are only 20 police division in Toronto.
+
+Significant outliers are evident in both the Mental Health Act (MHA) reports and the Major Crime Indicator data.
+These outliers may be attributable to the relatively small sample size, given that there are only 20 police divisions in Toronto.
+The limited number of divisions can lead to greater variability in the data, making it more susceptible to the influence of atypical observations.
 
@@
 ggplot(long_data, aes(x = Indicator, y = Count)) +
-  geom_boxplot(outlier.colour = "black", outlier.shape = 16, outlier.size = 2, fill = NA, color = "black") +
-  geom_jitter(width = 0.2, alpha = 0.5) +  
+  geom_boxplot(outlier.colour = "#F1948A", outlier.shape = 16, outlier.size = 2, fill = NA, color = "#A9CCE3") +
+  geom_jitter(width = 0.2, alpha = 0.5, color = "#F1948A") +  
   labs(title = "Boxplot of MHA Reports and Crimes",
@@
 
-## 7.2 Linear model
-The linear model indeed shows significant linear relationship between the number of MHA reports and major crime indicator.
-However, the adjusted R-squared in this model is 0.6388, indicating that there are 36.12% of the variance left unexplained by the model.
-This could imply that other factors, aside from MHA reports, might be influencing crime rates.
+## 7.2 Model accuracy
+
+The linear model reveals a significant linear relationship between the number of Mental Health Act (MHA) reports and the major crime indicator.
+However, the adjusted R-squared value of 0.6388 indicates that approximately 36.12% of the variance in crime rates remains unexplained by the model.
+This finding implies that, in addition to MHA reports, there are likely other influencing factors contributing to variations in crime rates.
 
@@
 #| label: Summary table
-#| fig-cap: Summary table of linear model
 #| echo: false
@@
   gt() %>%
-  tab_header(title = "Linear Model Summary") %>%
+  tab_header(title = "Linear Model Summary 2") %>%
   fmt_number(
@@
 
+\newpage
+
 ## 7.3 Model residual
-The residuals in this model are quite large, leading to a complex residual pattern, which may indicate potential issues with model fit or the presence of unaccounted variability in the data.
+
+The residuals in this model are relatively large, leading to a complex residual pattern.
+This complexity may indicate potential issues with model fit or the presence of unaccounted variability in the data.
 
@@
 
-plot(model, which=1)
+fitted_values <- fitted(model)
+residuals <- resid(model)
+
+ggplot(data = NULL, aes(x = fitted_values, y = residuals)) +
+  geom_point(color = "#A9CCE3") +     # Set point color to #A9CCE3
+  geom_smooth(method = "loess", color = "#F1948A", se = FALSE) +  # Smoothed line following residuals
+  labs(x = "Fitted Values", y = "Residuals") +  # Add axis labels
+  theme_minimal()
+
 ```
@@
 
-# 8.Reference
+# 8.Appendix
+
+```{=tex}
+\begin{table}[H]
+    \centering
+    \caption{Linear Model Summary 3}
+    \begin{tabular}{|l|l|}
+        \hline
+        \textbf{Statistic} & \textbf{Value} \\ \hline
+        Minimum & -10046.5 \\ \hline
+        1st Quartile (1Q) & -3273.8 \\ \hline
+        Median & -423.9 \\ \hline
+        3rd Quartile (3Q) & 4260.7 \\ \hline
+        Maximum & 6344.1 \\ \hline
+        \textbf{Coefficients} &  \\ \hline
+        (Intercept) & 6748.2438 \\ \hline
+        Std. Error (Intercept) & 2970.8029 \\ \hline
+        t value (Intercept) & 2.272 \\ \hline
+        Pr(>|t|) (Intercept) & 0.0373 * \\ \hline
+        num\_mha\_reports & 2.4509 \\ \hline
+        Std. Error (num\_mha\_reports) & 0.4397 \\ \hline
+        t value (num\_mha\_reports) & 5.574 \\ \hline
+        Pr(>|t|) (num\_mha\_reports) & 4.19e-05 *** \\ \hline
+        \textbf{Significance Codes} & 0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1 \\ \hline
+        \textbf{Residual Standard Error} & 4835 on 16 degrees of freedom \\ \hline
+        \textbf{Multiple R-squared} & 0.6601 \\ \hline
+        \textbf{Adjusted R-squared} & 0.6388 \\ \hline
+        \textbf{F-statistic} & 31.07 on 1 and 16 DF \\ \hline
+        \textbf{p-value of F-statistic} & 4.194e-05 \\ \hline
+    \end{tabular}
+\end{table}
+```
+
+```{=tex}
+\begin{table}[H]
+    \centering
+    \caption{Cleaned Data}
+    \begin{tabular}{|l|r|r|}
+        \hline
+        \textbf{Division} & \textbf{Number of Crimes} & \textbf{Number of MHA Reports} \\ \hline
+        D11 & 15894 & 5435 \\ \hline
+        D12 & 17151 & 5137 \\ \hline
+        D13 & 14180 & 4226 \\ \hline
+        D14 & 27590 & 12603 \\ \hline
+        D22 & 26578 & 6302 \\ \hline
+        D23 & 25782 & 5769 \\ \hline
+        D31 & 29099 & 6531 \\ \hline
+        D32 & 31273 & 7528 \\ \hline
+        D33 & 17836 & 5516 \\ \hline
+        D41 & 27591 & 6752 \\ \hline
+        D42 & 24037 & 5357 \\ \hline
+        D43 & 27509 & 7069 \\ \hline
+        D51 & 32666 & 10233 \\ \hline
+        D52 & 23487 & 5916 \\ \hline
+        D53 & 21199 & 6584 \\ \hline
+        D54 & 6347 & 1965 \\ \hline
+        D55 & 24524 & 8636 \\ \hline
+        NSA & 3992 & 755 \\ \hline
+    \end{tabular}
+\end{table}
+```
+
+\vspace{2em}
+\newpage
+
+# 9.Reference

### 2024-09-25 Paper/Paper.qmd
Add files via upload

@@
 
+The bar plot (Figure @fig-MHA) and the line plot (Figure @fig-MHA-2) show the number of MHA Apprehensions in each police division and its trend.
+
 \newpage
 
-## 3.1 Results of MHA Apprehensions Data
+## 3.1 Overview of MHA Apprehensions Data
 
 ```{r,fig.pos='H'}
-#| label: MHA Apprehensions Data
+#| label: fig-MHA
 #| fig-cap: Number of MHA reports in each Police division (2014-2024)
@@
 ```{r,fig.pos='H'}
-#| label: MHA Apprehensions Data 2
+#| label: fig-MHA-2
 #| fig-cap: Number of MHA reports in each Police division (2014-2024)
@@
 
+The bar plot (Figure @fig-MCI) and the line plot (Figure @fig-MCI-2) show the number of Major Crime Indicators in each police division and its trend.
+
 \newpage
 
-## 4.1 Results of Major Crime Indicator Data
+## 4.1 Overview of Major Crime Indicator Data
 
 ```{r,fig.pos='H'}
-#| label: Major Crime Indicator Data
+#| label: fig-MCI
 #| fig-cap: Number of Major Crime Indicator in each Police division (2014-2024)
@@
 ```{r,fig.pos='H'}
-#| label: Major Crime Indicator Data 2
+#| label: fig-MCI-2
 #| fig-cap: Number of Major Crime Indicator in each Police division (2014-2024)
@@
 
+## 5.1 Comparing two datasets
 ```{r,fig.pos='H'}
-#| label: comparison 
+#| label: fig-compare
 #| fig-cap: comparison between two data
@@
 ```{r,fig.pos='H'}
-#| label: comparison 2
+#| label: fig-compare-2
 #| fig-cap: comparison between two data
@@
 
-By comparing the two line plots,a positive relationship between the number of MHA reports and the number of Major Crime Indicator, the number of MHA reports seems to influence the number of crime incidents accross different police divisions.
-In order to confirm there are linear relationship between two factors, a linear model is made to fit in this situation based on two data:
+By comparing the two bar plots (Figure @fig-compare) and line plots (Figure @fig-compare-2),there seems to be a positive relationship between the number of MHA reports and the number of Major Crime Indicator.
+To check whether they are linearly related, a test for linear relationship is necessary.
+
+## 5.2 Introducing linear model
+In order to confirm there are linear relationship between two factors, a linear model is made to fit in this situation based on two data sets:
 
@@
 C = 6748.244 + 2.451 \times MHA 
-$$ in which:
+$$
+
+in which:
 
@@
 
-The corresponding graph of linear model is:
+The corresponding graph of linear model is @fig-linear-model
 
 ```{r,fig.pos='H'}
-#| label: Linear model
+#| label: fig-linear-model
 #| fig-cap: Linear Model
@@
 
-The p-values for both intercept and coefficient (MHA Reports) are below the significant level of 0.05, indicating a strong evidence that the relationship between MHA Reports and major crime indicator follows this linear model.
+The p-values for both intercept and coefficient (MHA Reports) are below the significant level of 0.05 (Figure @fig-Summary), indicating a strong evidence that the relationship between MHA Reports and major crime indicator follows this linear model.
 
 ```{r,fig.pos='H'}
-#| label: Linear model summary
+#| label: fig-Summary
+#| fig-cap: Linear Model Summary 1
 #| echo: false
@@
 Results shows that as the number of MHA reports rises, the major crime indicator also increases, based on data from 2014 to 2024.
-Furthermore, there appears to be a positive linear relationship between the number of MHA reports and the number of major crime indicator, which can be expressed as a linear model: C = 6748.244 + 2.451 \* MHA.
+Furthermore, there appears to be a positive linear relationship between the number of MHA reports and the number of major crime indicator, which can be expressed as a linear model $C = 6748.244 + 2.451 \times MHA$.
 
-The result of this paper shares the insight of the research by Public Safety Canada: "In the international research for both youth and adults analyzed for this report, findings suggest that certain mental health disorders, such as substance use disorder, conduct disorder, and antisocial personality disorder were shown to have a significant correlation to offending and re-offending." The p-value of our model is 4.19\*10\^-5, which is far below the significance level of 0.05, indicating a strong correlation between the number of Mental Health Act Apprehensions and the major crime indicator.[@PublicSafetyCanada]
+The result of this paper shares the insight of the research by Public Safety Canada: "In the international research for both youth and adults analyzed for this report, findings suggest that certain mental health disorders, such as substance use disorder, conduct disorder, and antisocial personality disorder were shown to have a significant correlation to offending and re-offending." The p-value of our model is $4.19 \times 10^{-5}$, which is far below the significance level of 0.05, indicating a strong correlation between the number of Mental Health Act Apprehensions and the major crime indicator.[@PublicSafetyCanada]
 
@@
 
-Significant outliers are evident in both the Mental Health Act (MHA) reports and the Major Crime Indicator data.
+Significant outliers are evident in both the Mental Health Act (MHA) reports and the Major Crime Indicator data (Figure @fig-Box-Plot).
 These outliers may be attributable to the relatively small sample size, given that there are only 20 police divisions in Toronto.
@@
 ```{r,fig.pos='H'}
-#| label: Box Plot
+#| label: fig-Box-Plot
 #| fig-cap: The box plot of two factors
@@
 The linear model reveals a significant linear relationship between the number of Mental Health Act (MHA) reports and the major crime indicator.
-However, the adjusted R-squared value of 0.6388 indicates that approximately 36.12% of the variance in crime rates remains unexplained by the model.
+However, the adjusted R-squared value of 0.6388 indicates that approximately 36.12% of the variance in crime rates remains unexplained by the model (Figure @fig-Summary-2).
 This finding implies that, in addition to MHA reports, there are likely other influencing factors contributing to variations in crime rates.
@@
 ```{r,fig.pos='H'}
-#| label: Summary table
+#| label: fig-Summary-2
+#| fig-cap: Linear Model Summary 2
 #| echo: false
@@
 
-The residuals in this model are relatively large, leading to a complex residual pattern.
+The residuals in this model are relatively large, leading to a complex residual pattern (Figure @fig-Residual-Plot).
 This complexity may indicate potential issues with model fit or the presence of unaccounted variability in the data.
@@
 ```{r,fig.pos='H'}
-#| label: Residual Plot
+#| label: fig-Residual-Plot
 #| fig-cap: The Residual Vs Fitted value plot
@@
 ```
-
 ```{=tex}
@@
 ```
-
+```{=tex}
 \vspace{2em}
 \newpage
-
+```
 # 9.Reference

### 2024-09-26 Paper/Paper.qmd
Final Draft of Paper
Adding citation for packages and equations. Enhancing graphs and table of contents.

@@
 
+## 2.1 Packages
 The data used in this paper came from the Open Data Toronto portal through the library opendatatoronto [@OpenDataToronto].
@@
 
+## 2.2 Clean data process
 The cleaned data comes from two raw data sets: Mental Health Act Apprehensions Dataset, which contains records of Mental Health Act (MHA) reports grouped by police division, and Major Crime Indicators Dataset, which includes records of crimes similarly categorized by police division.
@@
 
-$$
+\begin{equation}
 C = 6748.244 + 2.451 \times MHA 
-$$
+\label{eq:linear}
+\end{equation}
 
@@
 Results shows that as the number of MHA reports rises, the major crime indicator also increases, based on data from 2014 to 2024.
-Furthermore, there appears to be a positive linear relationship between the number of MHA reports and the number of major crime indicator, which can be expressed as a linear model $C = 6748.244 + 2.451 \times MHA$.
+Furthermore, there appears to be a positive linear relationship between the number of MHA reports and the number of major crime indicator, which can be expressed as a linear model $C = 6748.244 + 2.451 \times MHA$ \eqref{eq:linear}.
 

### 2024-09-26 scripts/00-simulated_data.R
Updating scripts
Deleted irrelevant coding and confirmed downloading paths.

@@ -1,35 +1,23 @@
-# Load necessary libraries
 library(dplyr)
-
-# Set seed for reproducibility
 set.seed(123)
 
-# Use the actual list of districts from your clean data
 districts <- c("D11", "D12", "D13", "D14", "D22", "D23", "D31", "D32", "D33", "D41", 
                "D42", "D43", "D51", "D52", "D53", "D54", "D55", "NSA")
 
-# Simulate the number of MHA reports based on the range you specified earlier
 num_mha_reports <- sample(1000:15000, length(districts), replace = TRUE)
 
-# Coefficients from the linear model
 intercept <- 6748.2438
 slope <- 2.4509
 
-# Calculate the expected number of crimes based on the linear model
 num_crimes <- intercept + slope * num_mha_reports
 
-# Add normally distributed noise to the number of crimes
-# Use the residual standard error from the model
 residual_std_error <- 4835
 num_crimes <- num_crimes + rnorm(length(districts), mean = 0, sd = residual_std_error)
 
-# Ensure the number of crimes is non-negative
 num_crimes <- pmax(num_crimes, 0)
 
-# Combine into a data frame
 simulated_data <- data.frame(District = districts, 
                              num_crimes = round(num_crimes), 
                              num_mha_reports = num_mha_reports)
 
-# Save the simulated data to a CSV file
 write.csv(simulated_data, "/home/rstudio/STA304_Paper_1_MaxLI/data/simulated_data/simulated_data.csv", row.names = FALSE)

### 2024-09-26 scripts/01-download_data.R

@@ -2,15 +2,12 @@
 library(dplyr)
 library(httr)
 
-# Define the URLs for the datasets
 mha_url <- "https://ckan0.cf.opendata.inter.prod-toronto.ca/dataset/mental-health-apprehensions/resource/4a92540b-6504-4b25-827d-4511b8b9f3d9/download/mental-health-apprehensions.csv"
 mci_url <- "https://ckan0.cf.opendata.inter.prod-toronto.ca/dataset/major-crime-indicators/resource/863bbdfb-a932-48b7-8a59-bf6585ef0563/download/major-crime-indicators.csv"
 
-# Download the datasets
 mha_data <- GET(mha_url)
 mci_data <- GET(mci_url)
 
-# Check if the download was successful
 if (status_code(mha_data) == 200) {
   write.csv(content(mha_data, "text"), 
             "/home/rstudio/STA304_Paper_1_MaxLI/data/raw_data/raw_data_mha.csv", 
@@ -27,5 +24,4 @@ if (status_code(mci_data) == 200) {
   stop("Failed to download MCI data")
 }
 
-# Print a message indicating that the data has been downloaded
 cat("Data downloaded successfully and saved to the specified locations.\n")

### 2024-09-26 scripts/02-data_cleaning.R

@@ -1,22 +1,16 @@
-# Load necessary libraries
 library(dplyr)
 
-# Read in the raw data
 mha_data <- read.csv("/home/rstudio/STA304_Paper_1_MaxLI/data/raw_data/raw_data_mha.csv")
 mci_data <- read.csv("/home/rstudio/STA304_Paper_1_MaxLI/data/raw_data/raw_data_mci.csv")
 
-# Summarize the number of MHA reports by district (using DIVISION as district)
 mha_summary <- mha_data %>%
   group_by(DIVISION) %>%
   summarise(num_mha_reports = n())
 
-# Summarize the number of crimes by district (using DIVISION as district)
 mci_summary <- mci_data %>%
   group_by(DIVISION) %>%
   summarise(num_crimes = n())
 
-# Merge the two summaries by District (DIVISION)
 clean_data <- merge(mci_summary, mha_summary, by = "DIVISION", all = TRUE)
 
-# Save the cleaned data to a CSV file
 write.csv(clean_data, "/home/rstudio/STA304_Paper_1_MaxLI/data/analysis_data/clean_data.csv", row.names = FALSE)

### 2024-09-26 scripts/03-test_data.R

@@ -1,27 +1,21 @@
-# Load necessary libraries
 library(dplyr)
 
-# Read in the cleaned data
 clean_data <- read.csv("/home/rstudio/STA304_Paper_1_MaxLI/data/analysis_data/clean_data.csv")
 
-# Basic structure and summary of the data
 cat("Structure of the cleaned data:\n")
 str(clean_data)
 
 cat("\nSummary of the cleaned data:\n")
 summary(clean_data)
 
-# Check for missing values
 cat("\nChecking for missing values:\n")
 missing_values <- sum(is.na(clean_data))
 cat("Total missing values:", missing_values, "\n")
 
-# Check unique districts
 cat("\nUnique districts in the dataset:\n")
 unique_districts <- unique(clean_data$DIVISION)
 print(unique_districts)
 
-# Basic statistics
 cat("\nBasic statistics for number of crimes and MHA reports:\n")
 crime_stats <- clean_data %>%
   summarise(
@@ -34,24 +28,16 @@ crime_stats <- clean_data %>%
   )
 print(crime_stats)
 
-# Visualize the data (optional, requires ggplot2)
-library(ggplot2)
-
-# Plot the number of crimes and MHA reports by district
-ggplot(clean_data, aes(x = reorder(DIVISION, -num_crimes), y = num_crimes)) +
-  geom_bar(stat = "identity", fill = "steelblue") +
-  labs(title = "Number of Crimes by District",
-       x = "District",
+library("ggplot2")
+model <- lm(num_crimes ~ num_mha_reports, data = clean_data)
+clean_data$predicted_crimes <- predict(model, newdata = clean_data)
+ggplot(clean_data, aes(x = num_mha_reports, y = num_crimes)) +
+  geom_point(color = "#A9CCE3") +  
+  geom_line(aes(y = predicted_crimes), color = "#F1948A") + 
+  labs(title = "Number of Crimes vs. Number of MHA Reports",
+       x = "Number of MHA Reports",
        y = "Number of Crimes") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+  theme_minimal()
 
-ggplot(clean_data, aes(x = reorder(DIVISION, -num_mha_reports), y = num_mha_reports)) +
-  geom_bar(stat = "identity", fill = "salmon") +
-  labs(title = "Number of MHA Reports by District",
-       x = "District",
-       y = "Number of MHA Reports") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
[…]

### 2024-09-26 scripts/04-model.R

@@ -1,45 +1,26 @@
-# Load necessary libraries
 library(dplyr)
 library(ggplot2)
 
-# Read in the cleaned data
 clean_data <- read.csv("/home/rstudio/STA304_Paper_1_MaxLI/data/analysis_data/clean_data.csv")
 
-# Fit a linear regression model
 model <- lm(num_crimes ~ num_mha_reports, data = clean_data)
 
-# Summary of the model
 cat("Model Summary:\n")
 summary(model)
 
-# Check residuals
 par(mfrow = c(2, 2))
 plot(model)
 
-# Predicting number of crimes based on MHA reports
 clean_data$predicted_crimes <- predict(model, newdata = clean_data)
 
-# Visualizing the relationship
-ggplot(clean_data, aes(x = num_mha_reports, y = num_crimes)) +
-  geom_point(color = "blue") +
-  geom_line(aes(y = predicted_crimes), color = "red") +
-  labs(title = "Number of Crimes vs. Number of MHA Reports",
-       x = "Number of MHA Reports",
-       y = "Number of Crimes") +
-  theme_minimal()
-
-# Create directory for model if it doesn't exist
 if (!dir.exists("/home/rstudio/STA304_Paper_1_MaxLI/model/")) {
   dir.create("/home/rstudio/STA304_Paper_1_MaxLI/model/", recursive = TRUE)
 }
 
-# Save the model as an RDS file
 saveRDS(model, "/home/rstudio/STA304_Paper_1_MaxLI/model/crime_model.rds")
 
-# Save the model summary to a text file
 sink("/home/rstudio/STA304_Paper_1_MaxLI/model/model_summary.txt")
 summary(model)
 sink()
 
-# Print a message indicating that the modeling is complete
 cat("\nModeling complete. Check the output above for details.\n")

### 2024-09-26 scripts/05-replications.R

@@ -1,44 +1,23 @@
-# Load necessary libraries
 library(dplyr)
 library(ggplot2)
 
-# Read in the cleaned data
 clean_data <- read.csv("/home/rstudio/STA304_Paper_1_MaxLI/data/analysis_data/clean_data.csv")
-
-# Load the saved model
 model <- readRDS("/home/rstudio/STA304_Paper_1_MaxLI/model/crime_model.rds")
-
-# Generate predictions using the loaded model and round them to the nearest integer
 clean_data$predicted_crimes <- round(predict(model, newdata = clean_data))
 
-# Check the first few rows of the data with predictions
 cat("First few rows of the cleaned data with predictions:\n")
 head(clean_data)
 
-# Visualizing the predicted vs actual number of crimes
-ggplot(clean_data, aes(x = num_mha_reports)) +
-  geom_point(aes(y = num_crimes), color = "blue", alpha = 0.6, size = 2) +
-  geom_line(aes(y = predicted_crimes), color = "red") +
-  labs(title = "Actual vs Predicted Number of Crimes",
-       x = "Number of MHA Reports",
-       y = "Number of Crimes") +
-  theme_minimal()
-
-# Calculate residuals
 clean_data$residuals <- clean_data$num_crimes - clean_data$predicted_crimes
 
-# Summary of residuals
 cat("\nSummary of residuals:\n")
 summary(clean_data$residuals)
 
-# Create directory for replication results if it doesn't exist
 if (!dir.exists("/home/rstudio/STA304_Paper_1_MaxLI/data/replication/")) {
   dir.create("/home/rstudio/STA304_Paper_1_MaxLI/data/replication/", recursive = TRUE)
 }
 
-# Save the results with predictions to a CSV file
 write.csv(clean_data, "/home/rstudio/STA304_Paper_1_MaxLI/data/replication/clean_data_with_predictions.csv", row.names = FALSE)
 
-# Print a message indicating that the replication is complete
 cat("\nReplication complete. Check the output above for details.\n")
 
